\documentclass[12pt]{article}

\usepackage{amsmath, amssymb}
\usepackage{geometry}
\usepackage{setspace}
\usepackage{titlesec}
\usepackage{lmodern}
\usepackage{xcolor}
\usepackage{enumitem}

\geometry{margin=1in}
\setstretch{1.2}
\titleformat{\section}{\normalfont\Large\bfseries}{\thesection}{1em}{}
\titleformat{\subsection}{\normalfont\large\bfseries}{\thesubsection}{1em}{}
\pagenumbering{gobble}

\begin{document}

\begin{center}
    \LARGE \textbf{Unit 3: Ratios, Rates, and Percents} \\[6pt]
    \Large \textbf{Topic 2: Unit Rate and Proportional Relationships}
\end{center}

\vspace{1em}

\section*{Concept Summary}

A relationship between two quantities is \textbf{proportional} if one quantity is a constant multiple of the other.  
Algebraically: \(y = kx\) where \(k\) is the \textbf{constant of proportionality} or \textbf{unit rate}.  
Key features:
\begin{itemize}
  \item Graph passes through the origin \((0,0)\).
  \item Slope equals \(k\). Each 1 unit increase in \(x\) changes \(y\) by \(k\).
  \item From a table, \(k = \dfrac{y}{x}\) is the same for all nonzero \(x\).
\end{itemize}

A relationship is \textbf{not proportional} if it has a nonzero intercept or if \(\dfrac{y}{x}\) is not constant. Any form \(y = kx + b\) with \(b \ne 0\) is not proportional.

\section*{Core Skills}
\begin{itemize}
  \item Compute a \textbf{unit rate} by dividing per one unit of the denominator.
  \item Test proportionality using equal ratios or a straight line through the origin.
  \item Write and use \(y = kx\) to predict unknown values and interpret slope as rate.
  \item Convert among tables, equations, and graphs for the same proportional model.
\end{itemize}

\section*{Example 1: Finding a Unit Rate}

A runner travels 9 miles in 72 minutes. Find miles per minute and miles per hour.

\textbf{Miles per minute:} \(\dfrac{9}{72} = \dfrac{1}{8} = 0.125\) mi per min.  
\textbf{Miles per hour:} multiply by 60: \(0.125 \times 60 = 7.5\) mph.  
\textbf{Model:} \(d = 7.5t\) where \(t\) is hours.

\section*{Example 2: Proportional or Not}

Decide if the relationship is proportional: \(y = 4x - 3\).

The intercept is \(-3\). Since the graph does not pass through \((0,0)\), this is not proportional.

\section*{Example 3: From Table to Equation}

\[
\begin{array}{c|cccc}
x & 2 & 5 & 8 & 10\\ \hline
y & 6 & 15 & 24 & 30
\end{array}
\]
Compute \(\dfrac{y}{x}\): \(\dfrac{6}{2}=3,\ \dfrac{15}{5}=3,\ \dfrac{24}{8}=3,\ \dfrac{30}{10}=3\). Constant ratio 3.  
\textbf{Equation:} \(y = 3x\). Here \(k=3\).

\section*{Example 4: Graph Interpretation}

A graph of a line through \((0,0)\) and \((4,10)\) represents \(y\) versus \(x\).  
Slope \(k = \dfrac{10-0}{4-0} = \dfrac{10}{4} = 2.5\).  
\textbf{Equation:} \(y = 2.5x\). The unit rate is 2.5 units of \(y\) per 1 unit of \(x\).

\section*{Example 5: Context Prediction with \(y = kx\)}

A fruit stand sells apples at a constant price. If 3 pounds cost \$7.50, find the price for 8 pounds.

Unit rate \(k = \dfrac{7.50}{3} = 2.50\) dollars per pound.  
Model \(C = 2.5x\). For \(x=8\), \(C = 2.5 \cdot 8 = \$20\).

\section*{Key Takeaways}
\begin{itemize}
  \item Proportional relationships have the form \(y = kx\) and pass through the origin.
  \item The constant \(k\) is the unit rate and equals the slope of the graph.
  \item Test proportionality by checking equal ratios or intercept at the origin.
  \item Use \(y = kx\) to convert, compare, and predict values in context.
\end{itemize}

\newpage

% ============================================================
% QUESTIONS — TOPIC 2: UNIT RATE AND PROPORTIONAL RELATIONSHIPS
% ============================================================

\section*{Practice Questions: Unit Rate and Proportional Relationships}

\subsection*{Part A: Finding Unit Rates}
Find the unit rate for each situation.
\begin{enumerate}
  \item A car travels 300 miles in 5 hours.
  \item 8 notebooks cost \$20.
  \item 240 miles on 8 gallons of gas.
  \item 15 meters of fabric costs \$90.
  \item 120 dollars earned for 15 hours of work.
\end{enumerate}

\subsection*{Part B: Testing for Proportionality}
Decide whether each relationship is proportional. Explain briefly.
\begin{enumerate}
  \setcounter{enumi}{5}
  \item \(y = 2.5x\)
  \item \(y = 4x + 1\)
  \item \(y = \dfrac{3}{x}\)
  \item \(y = 7x - 2\)
  \item Table:  
  \[
  \begin{array}{c|cccc}
  x & 1 & 2 & 3 & 4\\ \hline
  y & 5 & 10 & 15 & 20
  \end{array}
  \]
\end{enumerate}

\subsection*{Part C: Writing Equations for Proportional Relationships}
\begin{enumerate}
  \setcounter{enumi}{10}
  \item A car travels 180 miles in 3 hours. Write an equation for miles \(y\) as a function of hours \(x\).
  \item 5 pounds of apples cost \$12.50. Write an equation for total cost \(C\) in dollars for \(p\) pounds.
  \item The line through \((0,0)\) and \((6,18)\) models \(y\) versus \(x\). Write the equation.
  \item For each pair \((x,y)\), check if the data are proportional:  
  \((2,6), (5,15), (8,24), (10,30)\).
  \item A constant-speed train covers 240 miles in 4 hours. Write and interpret the proportional equation.
\end{enumerate}

\subsection*{Part D: Using Proportional Models}
\begin{enumerate}
  \setcounter{enumi}{15}
  \item A recipe uses 4 cups of flour for 10 servings. How many cups are needed for 25 servings?
  \item A photo is 4 inches wide and 6 inches tall. A larger version keeps the same proportions and is 10 inches wide. Find the new height.
  \item A bicycle moves 60 miles in 3 hours. How long will it take to go 100 miles at the same rate?
  \item At a store, 7 pounds of oranges cost \$10.50. Find the cost of 4 pounds.
  \item The distance a car travels is proportional to time. If 5 hours corresponds to 350 miles, how far in 8 hours?
\end{enumerate}

\subsection*{Part E: SAT-Style Applications}
\begin{enumerate}
  \setcounter{enumi}{20}
  \item A printer produces 180 pages in 12 minutes. How long will it take to print 525 pages at the same rate?
  \item If \(y\) varies directly with \(x\) and \(y=12\) when \(x=3\), find \(y\) when \(x=8\).
  \item A faucet fills a sink with water at a constant rate. It fills 2 gallons in 1.5 minutes. How many gallons in 10 minutes?
  \item The graph of \(y\) vs. \(x\) passes through \((0,0)\) and \((9,6)\). Write the equation and find \(y\) when \(x=12\).
  \item Two variables \(x\) and \(y\) are proportional. When \(x=14\), \(y=49\). Write the equation and find \(y\) when \(x=20\).
\end{enumerate}

\newpage

% ============================================================
% SOLUTIONS — UNIT 3, TOPIC 2: UNIT RATE AND PROPORTIONAL RELATIONSHIPS
% ============================================================

\section*{Answer Key and Solutions: Unit Rate and Proportional Relationships}

\subsection*{Part A Solutions: Finding Unit Rates}
\begin{enumerate}
  \item \(300\text{ mi}/5\text{ h} = \boxed{60\text{ mph}}\)
  \item \(\$20/8 = \boxed{\$2.50 \text{ per notebook}}\)
  \item \(240\text{ mi}/8\text{ gal} = \boxed{30\text{ mpg}}\)
  \item \(\$90/15\text{ m} = \boxed{\$6 \text{ per meter}}\)
  \item \(\$120/15\text{ h} = \boxed{\$8 \text{ per hour}}\)
\end{enumerate}

\subsection*{Part B Solutions: Testing for Proportionality}
\begin{enumerate}
  \setcounter{enumi}{5}
  \item \(y=2.5x\). Passes through origin. \(\boxed{\text{Proportional}}\)
  \item \(y=4x+1\). Intercept \(1\neq0\). \(\boxed{\text{Not proportional}}\)
  \item \(y=\dfrac{3}{x}\). Ratio \(y/x\) not constant. \(\boxed{\text{Not proportional}}\)
  \item \(y=7x-2\). Intercept \(-2\neq0\). \(\boxed{\text{Not proportional}}\)
  \item Table: \(y/x=5\) for all \(x\). \(\boxed{\text{Proportional with } y=5x}\)
\end{enumerate}

\subsection*{Part C Solutions: Writing Equations}
\begin{enumerate}
  \setcounter{enumi}{10}
  \item Speed \(=180/3=60\). \(\boxed{y=60x}\)
  \item Unit price \(=12.50/5=2.50\). \(\boxed{C=2.5p}\)
  \item Slope \(=18/6=3\). \(\boxed{y=3x}\)
  \item Ratios \(6/2=15/5=24/8=30/10=3\). \(\boxed{\text{Proportional, } y=3x}\)
  \item Speed \(=240/4=60\). \(\boxed{y=60x}\). Interpretation: 60 miles per hour.
\end{enumerate}

\subsection*{Part D Solutions: Using Proportional Models}
\begin{enumerate}
  \setcounter{enumi}{15}
  \item Cups per serving \(=4/10=0.4\). For 25: \(0.4\cdot25=\boxed{10\text{ cups}}\)
  \item Scale factor \(=10/4=2.5\). Height \(=6\cdot2.5=\boxed{15\text{ in}}\)
  \item Speed \(=60/3=20\). Time \(=100/20=\boxed{5\text{ h}}\)
  \item Unit price \(=10.50/7=1.50\). Cost \(=1.50\cdot4=\boxed{\$6}\)
  \item Speed \(=350/5=70\). Distance \(=70\cdot8=\boxed{560\text{ miles}}\)
\end{enumerate}

\subsection*{Part E Solutions: SAT-Style Applications}
\begin{enumerate}
  \setcounter{enumi}{20}
  \item Rate \(=180/12=15\) pages per min. Time \(=525/15=\boxed{35\text{ min}}\)
  \item \(y=kx\). \(12=3k \Rightarrow k=4\). For \(x=8\): \(\boxed{y=32}\)
  \item Rate \(=2/1.5=\tfrac{4}{3}\) gal per min. In 10 min: \(10\cdot\tfrac{4}{3}=\boxed{\tfrac{40}{3}\text{ gal}}\)
  \item Slope \(=6/9=2/3\). \(\boxed{y=\tfrac{2}{3}x}\). For \(x=12\): \(\boxed{y=8}\)
  \item \(y=kx\). \(49=14k \Rightarrow k=3.5\). For \(x=20\): \(\boxed{y=70}\)
\end{enumerate}


\end{document}
