\documentclass[14pt]{extarticle}

% ---------- PACKAGES ----------
\usepackage{amsmath, amssymb}
\usepackage{geometry}
\usepackage{setspace}
\usepackage{titlesec}
\usepackage{xcolor}
\usepackage{helvet}
\usepackage{enumitem}

% ---------- PAGE SETUP ----------
\geometry{margin=1in}
\setstretch{1.5}
\setlength{\parindent}{0pt}
\setlength{\parskip}{0.75em}
\setlist{itemsep=0.75\baselineskip, topsep=0.5\baselineskip}
\titleformat{\section}{\normalfont\Large\bfseries}{\thesection}{1em}{}
\titleformat{\subsection}{\normalfont\large\bfseries}{\thesubsection}{1em}{}
\renewcommand{\familydefault}{\sfdefault}
\renewcommand{\emph}[1]{\textbf{#1}}

% ---------- DOCUMENT ----------
\begin{document}
\raggedright
\pagenumbering{gobble}

\begin{center}
    \LARGE \textbf{Unit 1: Linear Relationships and Equations} \\[6pt]
    \Large \textbf{Topic 7: Graphing and Interpreting Linear Models}
\end{center}

\vspace{1em}

\section*{Concept Summary}

Every linear equation represents a straight line on the coordinate plane.  
Understanding how the numbers in an equation affect its graph is key to solving many SAT questions.

\subsection*{Slope–Intercept Form and Graphing}
For a line written as
\[
y = mx + b
\]
\begin{itemize}
    \item \(m\) is the \textbf{slope}, showing how steep the line is.
    \item \(b\) is the \textbf{\(y\)-intercept}, showing where the line crosses the \(y\)-axis.
\end{itemize}

To graph:
1. Plot the \(y\)-intercept \((0, b)\).
2. Use the slope to find another point:  
   rise = change in \(y\), run = change in \(x\).
3. Connect the points with a straight line.

\textbf{Example:}  
For \(y = 2x + 1\):
- \(m = 2\): rise 2, run 1  
- \(b = 1\): start at \((0, 1)\)

\subsection*{Standard Form and Intercepts}
A line in standard form
\[
Ax + By = C
\]
can be graphed quickly using intercepts:
\[
\text{\(x\)-intercept: set } y = 0, \quad \text{\(y\)-intercept: set } x = 0.
\]
\textbf{Example:}  
For \(2x + 3y = 12\):
\[
x\text{-int: } (6, 0), \quad y\text{-int: } (0, 4)
\]

\subsection*{Interpreting Linear Models}
Linear equations often model real-world relationships, such as cost, distance, or temperature over time.

\begin{itemize}
    \item The \textbf{slope} represents the \textbf{rate of change} (how much one quantity changes per unit of another).
    \item The \textbf{\(y\)-intercept} represents the \textbf{starting value} when \(x = 0\).
\end{itemize}

\textbf{Example:}  
A taxi fare is given by \(C = 2.5x + 4\), where \(C\) is cost and \(x\) is miles.
\[
\text{slope } 2.5 \Rightarrow \text{cost per mile}, \quad \text{intercept } 4 \Rightarrow \text{base fee}.
\]

\subsection*{Intersections of Lines}
The intersection point of two lines represents the \textbf{solution to both equations}.  
It can be found by solving the system of equations simultaneously.

\textbf{Example:}
\[
\begin{cases}
y = 2x + 3 \\
y = -x + 9
\end{cases}
\]
Set equal:
\[
2x + 3 = -x + 9 \Rightarrow 3x = 6 \Rightarrow x = 2, \quad y = 7
\]
\[
\boxed{(2, 7)}
\]
Interpretation: both models have the same value at \(x = 2\).

\section*{Key Takeaways}
\begin{itemize}
    \item In \(y = mx + b\), \(m\) controls steepness and \(b\) controls vertical position.
    \item The intersection of two lines represents the solution to a system.
    \item On the SAT, slope often represents a rate (like cost per mile or speed).
    \item Understand how changes in \(m\) and \(b\) shift the graph:
    \begin{itemize}
        \item Increasing \(m\): steeper line
        \item Increasing \(b\): line shifts upward
    \end{itemize}
\end{itemize}

\newpage

% ============================================================
% QUESTIONS — TOPIC 7: GRAPHING AND INTERPRETING LINEAR MODELS
% ============================================================

\section*{Practice Questions: Graphing and Interpreting Linear Models}

\subsection*{Part A: Slope and Intercept Identification}
\begin{enumerate}
    \item Identify the slope and \(y\)-intercept of \(y = 3x - 4\).
    \item Identify the slope and \(y\)-intercept of \(y = -\tfrac{1}{2}x + 6\).
    \item Identify the slope and \(y\)-intercept of \(y = 5 - 2x\).
    \item Write the slope and \(y\)-intercept of \(2y = 8x + 10\).
    \item Which of the following lines is steeper: \(y = 4x + 1\) or \(y = 2x + 3\)?
\end{enumerate}

\subsection*{Part B: Finding and Using Intercepts}
\begin{enumerate}
    \setcounter{enumi}{5}
    \item Find the \(x\)- and \(y\)-intercepts of \(2x + y = 8\).
    \item Find the \(x\)- and \(y\)-intercepts of \(3x + 6y = 12\).
    \item Find the \(x\)-intercept of \(4x - 2y = 12\).
    \item Find the \(y\)-intercept of \(5x + 10y = 20\).
    \item Determine whether the line \(4x + 2y = 8\) passes through the point \((0,4)\).
\end{enumerate}

\subsection*{Part C: Effects of Slope and Intercept Changes}
\begin{enumerate}
    \setcounter{enumi}{10}
    \item Compare the graphs of \(y = 2x + 3\) and \(y = 2x - 1\).  
          How are they different?
    \item Compare the graphs of \(y = 3x + 2\) and \(y = -3x + 2\).  
          How are they different?
    \item How does increasing \(m\) in \(y = mx + b\) affect the steepness of the line?
    \item How does increasing \(b\) in \(y = mx + b\) affect the graph’s position?
    \item The line \(y = 4x + 2\) is shifted downward by 5 units.  
          Write the new equation.
\end{enumerate}

\subsection*{Part D: Intersections and Systems}
\begin{enumerate}
    \setcounter{enumi}{15}
    \item Find the intersection of \(y = x + 3\) and \(y = -x + 7\).
    \item Find the intersection of \(y = 2x + 4\) and \(y = -x + 1\).
    \item The lines \(y = 3x - 2\) and \(y = 3x + 5\) — do they intersect? Explain.
    \item The lines \(y = -2x + 8\) and \(2x + y = 4\) — find their intersection point.
    \item Find the \(x\)-value where \(y = -x + 10\) and \(y = 2x + 1\) have the same output.
\end{enumerate}

\subsection*{Part E: SAT-Style Applications}
\begin{enumerate}
    \setcounter{enumi}{20}
    \item A cell phone plan costs \$40 per month plus \$10 per gigabyte of data.  
          Write the equation for the total cost \(C\) in terms of data \(g\).
    \item A car’s value decreases from \$20,000 to \$14,000 over 3 years.  
          Write a linear equation for value \(V\) as a function of years \(t\).
    \item The temperature is 50°F at 6 AM and 62°F at noon.  
          Write the linear model for temperature \(T\) as a function of time \(t\) (hours after 6 AM).
    \item Two water tanks fill at different rates.  
          Tank A: \(y = 4x + 10\); Tank B: \(y = 6x + 2\).  
          Find when both have the same amount of water.
    \item A line passes through the origin with slope \(-\tfrac{3}{4}\).  
          Write its equation and describe the direction of the line.
\end{enumerate}

\newpage

% ============================================================
% SOLUTIONS — TOPIC 7: GRAPHING AND INTERPRETING LINEAR MODELS
% ============================================================

\section*{Answer Key and Solutions: Graphing and Interpreting Linear Models}

\subsection*{Part A Solutions: Slope and Intercept Identification}
\begin{enumerate}
    \item \(y=3x-4:\ m=\boxed{3},\ b=\boxed{-4}\).
    \item \(y=-\tfrac12 x+6:\ m=\boxed{-\tfrac12},\ b=\boxed{6}\).
    \item \(y=5-2x=-2x+5:\ m=\boxed{-2},\ b=\boxed{5}\).
    \item \(2y=8x+10\Rightarrow y=4x+5:\ m=\boxed{4},\ b=\boxed{5}\).
    \item Compare \(|m|\): \(4\) vs \(2\). Steeper: \(\boxed{y=4x+1}\).
\end{enumerate}

\subsection*{Part B Solutions: Finding and Using Intercepts}
\begin{enumerate}
    \setcounter{enumi}{5}
    \item \(2x+y=8:\ x\text{-int }(4,0),\ y\text{-int }(0,8)\).
    \item \(3x+6y=12:\ x\text{-int }(4,0),\ y\text{-int }(0,2)\).
    \item \(4x-2y=12:\) set \(y=0\Rightarrow 4x=12\Rightarrow x=3\Rightarrow \boxed{(3,0)}\).
    \item \(5x+10y=20:\) set \(x=0\Rightarrow 10y=20\Rightarrow y=2\Rightarrow \boxed{(0,2)}\).
    \item \(4(0)+2(4)=8\) is true \(\Rightarrow\) \boxed{Yes, it passes through }(0,4).
\end{enumerate}

\subsection*{Part C Solutions: Effects of Slope and Intercept Changes}
\begin{enumerate}
    \setcounter{enumi}{10}
    \item Same slope \(m=2\Rightarrow\) parallel; \(y\)-intercepts differ by \(4\). \(y=2x-1\) is \(\boxed{4}\) units below \(y=2x+3\).
    \item Same \(y\)-intercept \(2\); slopes \(3\) and \(-3\): one rises, one falls; \(\boxed{\text{not perpendicular}}\) (product \(-9\)).
    \item Increasing \(|m|\) makes the line \(\boxed{\text{steeper}}\).
    \item Increasing \(b\) shifts the line \(\boxed{\text{up}}\) (vertical translation).
    \item Down by \(5:\ y=4x+2-5=\boxed{y=4x-3}\).
\end{enumerate}

\subsection*{Part D Solutions: Intersections and Systems}
\begin{enumerate}
    \setcounter{enumi}{15}
    \item \(x+3=-x+7\Rightarrow 2x=4\Rightarrow x=2,\ y=5\Rightarrow \boxed{(2,5)}\).
    \item \(2x+4=-x+1\Rightarrow 3x=-3\Rightarrow x=-1,\ y=2\Rightarrow \boxed{(-1,2)}\).
    \item Same slope \(m=3\), different intercepts \(\Rightarrow\) \boxed{parallel, no intersection}.
    \item \(y=-2x+8\) and \(y=4-2x\): setting equal gives \(8=4\) (false) \(\Rightarrow\) \boxed{no solution (parallel)}.
    \item \(-x+10=2x+1\Rightarrow 3x=9\Rightarrow \boxed{x=3}\).
\end{enumerate}

\subsection*{Part E Solutions: SAT-Style Applications}
\begin{enumerate}
    \setcounter{enumi}{20}
    \item Cost model: \(\boxed{C=10g+40}\) (slope = \$10/GB, intercept = \$40).
    \item Slope \(\frac{14000-20000}{3}=-2000\); initial \(20000\Rightarrow \boxed{V=-2000t+20000}\).
    \item Points \((0,50)\) and \((6,62)\): \(m=\frac{62-50}{6}=2\Rightarrow \boxed{T=2t+50}\).
    \item Solve \(4x+10=6x+2\Rightarrow -2x=-8\Rightarrow x=4\). Amount \(y=4(4)+10=26\Rightarrow \boxed{x=4,\ y=26}\).
    \item Through origin, slope \(-\tfrac34\Rightarrow \boxed{y=-\tfrac34 x}\). The line \(\boxed{\text{decreases}}:\) down \(3\) for right \(4\).
\end{enumerate}

\newpage

% ============================================================
% QUESTIONS — TOPIC 3: INTERPRETING INTERSECTIONS AND SOLUTION TYPES
% ============================================================

\section*{Practice Questions: Interpreting Intersections and Solution Types}

\subsection*{Part A: Identify Solution Type by Inspection}
\begin{enumerate}
    \item \(\begin{cases} y = 2x + 1 \\ y = -x + 4 \end{cases}\)
    \item \(\begin{cases} y = 3x + 5 \\ y = 3x - 2 \end{cases}\)
    \item \(\begin{cases} 2x + 4y = 8 \\ x + 2y = 4 \end{cases}\)
    \item \(\begin{cases} y = -\tfrac{1}{2}x + 3 \\ y = 2x + 1 \end{cases}\)
    \item \(\begin{cases} 3x - 2y = 6 \\ 6x - 4y = 12 \end{cases}\)
\end{enumerate}

\subsection*{Part B: Determine Solution Type Algebraically}
\begin{enumerate}
    \setcounter{enumi}{5}
    \item \(\begin{cases} 2x + y = 10 \\ 4x + 2y = 20 \end{cases}\)
    \item \(\begin{cases} 3x - 2y = 5 \\ 6x - 4y = 9 \end{cases}\)
    \item \(\begin{cases} x + y = 8 \\ 2x + 2y = 18 \end{cases}\)
    \item \(\begin{cases} 4x + y = 12 \\ 8x + 2y = 26 \end{cases}\)
    \item \(\begin{cases} 5x - 3y = 7 \\ 10x - 6y = 14 \end{cases}\)
\end{enumerate}

\subsection*{Part C: Solve and Interpret the Intersection Point}
\begin{enumerate}
    \setcounter{enumi}{10}
    \item \(\begin{cases} y = 2x + 5 \\ y = -x + 8 \end{cases}\)
    \item \(\begin{cases} 3x + y = 12 \\ 2x - y = 3 \end{cases}\)
    \item \(\begin{cases} 4x - y = 11 \\ x + y = 4 \end{cases}\)
    \item \(\begin{cases} y = -2x + 6 \\ y = x - 3 \end{cases}\)
    \item \(\begin{cases} 2x + 3y = 12 \\ 3x + 2y = 12 \end{cases}\)
\end{enumerate}

\subsection*{Part D: Word Problems and Interpretation}
\begin{enumerate}
    \setcounter{enumi}{15}
    \item A gym charges \$30 per month plus \$5 per class. Another gym charges \$10 per month plus \$8 per class. At what number of classes will the cost be the same, and what will the cost be?
    \item One phone plan charges \$50 plus \$0.10 per minute, while another charges \$20 plus \$0.25 per minute. Find the number of minutes where both cost the same and what the total cost is.
    \item Two taxi services charge as follows:  
    Service A: \$3 plus \$1.50 per mile  
    Service B: \$6 plus \$1.00 per mile  
    After how many miles will both cost the same?
    \item The total cost of 3 pens and 2 notebooks is \$11. The total cost of 6 pens and 4 notebooks is \$22. How many solutions does this system have, and why?
    \item Two linear equations represent the same budget line in an economics problem. What kind of system is this, and what does that mean about the feasible combinations?
\end{enumerate}

% ============================================================
% SOLUTIONS — TOPIC 3: INTERPRETING INTERSECTIONS AND SOLUTION TYPES
% ============================================================

\section*{Answer Key and Solutions: Interpreting Intersections and Solution Types}

\subsection*{Part A Solutions: Identify by Inspection}
\begin{enumerate}
  \item Slopes 2 and -1. Different slopes → one solution.
  \item Slopes both 3. Different intercepts → no solution.
  \item \(2x+4y=8\) is \(2(x+2y=4)\). Same line → infinitely many solutions.
  \item Slopes \(-\tfrac12\) and 2. Different slopes → one solution.
  \item \(6x-4y=2(3x-2y)\). Same line → infinitely many solutions.
\end{enumerate}

\subsection*{Part B Solutions: Determine Algebraically}
\begin{enumerate}
  \setcounter{enumi}{5}
  \item Second is 2 times the first. Identity \(0=0\) after elimination → infinitely many.
  \item Left sides proportional, constants not. Contradiction \(0=c\) → no solution.
  \item \(2(x+y)=16 \neq 18\). Contradiction → no solution.
  \item Doubling \(4x+y=12\) gives \(8x+2y=24 \neq 26\). Contradiction → no solution.
  \item Second is 2 times the first. Identity → infinitely many.
\end{enumerate}

\subsection*{Part C Solutions: Solve and Interpret}
\begin{enumerate}
  \setcounter{enumi}{10}
  \item Set equal: \(2x+5=-x+8 \Rightarrow 3x=3 \Rightarrow x=1\). \(y=2(1)+5=7\). \(\boxed{(1,7)}\)
  \item Add: \((3x+y)+(2x-y)=12+3 \Rightarrow 5x=15 \Rightarrow x=3\). Then \(y=12-9=3\). \(\boxed{(3,3)}\)
  \item From \(x+y=4\) get \(y=4-x\). Substitute: \(4x-(4-x)=11 \Rightarrow 5x=15 \Rightarrow x=3\). \(y=1\). \(\boxed{(3,1)}\)
  \item Set equal: \(-2x+6=x-3 \Rightarrow -3x=-9 \Rightarrow x=3\). \(y=0\). \(\boxed{(3,0)}\)
  \item Eliminate: \(6x+9y=36\) and \(6x+4y=24\). Subtract → \(5y=12 \Rightarrow y=\tfrac{12}{5}\). Then \(2x+3(\tfrac{12}{5})=12 \Rightarrow 2x=\tfrac{24}{5} \Rightarrow x=\tfrac{12}{5}\). \(\boxed{(\tfrac{12}{5}, \tfrac{12}{5})}\)
\end{enumerate}

\subsection*{Part D Solutions: Word Problems}
\begin{enumerate}
  \setcounter{enumi}{15}
  \item Costs equal when \(30+5c=10+8c \Rightarrow 20=3c \Rightarrow c=\tfrac{20}{3}\). Cost \(=30+5(\tfrac{20}{3})=\tfrac{190}{3}\). Note: non-integer class count.
  \item \(50+0.10m=20+0.25m \Rightarrow 30=0.15m \Rightarrow m=200\). Cost \(=70\) dollars.
  \item \(3+1.5x=6+1.0x \Rightarrow 0.5x=3 \Rightarrow x=6\) miles.
  \item \(3p+2n=11\) and \(6p+4n=22\) are dependent. Infinitely many solutions. Not enough info for unique prices.
  \item Same budget line means a dependent system. Infinitely many feasible combinations along that line subject to any additional constraints.
\end{enumerate}



\end{document}
